---
title: "HD Code"
---


```{r}
# load libraries
library(ggplot2)
library(DESeq2)
library(org.Hs.eg.db)
```



#Data retrieval part
```{r}
HDData <- read.csv("data/HD_Counts.csv", header=T, check.names=FALSE, row.names=1)
HDmeta <- read.csv("data/HD_Meta.csv", header=T, check.names = FALSE,  row.names=1)
```


#Check step
```{r}

### Check that sample names match in both files
all(colnames(HDData) %in% rownames(HDmeta))
all(colnames(HDData) == rownames(HDmeta))

```



#DESeq Object for comparing HD VS. Control
```{r}
## Create DESeq2Dataset object
HDdds <- DESeqDataSetFromMatrix(countData = HDData, colData = HDmeta, design = ~ SampleType)
#View(counts(HDdds))
## Run analysis
HDdds <- DESeq(HDdds)
```



```{r}
print(HDdds)

```



#generating normalized counts
```{r}
HDdds2 <- estimateSizeFactors(HDdds)
HDnormalized_counts <- counts(HDdds2, normalized=TRUE)
write.csv(HDnormalized_counts, file="data/HDnormalized_counts.csv")

```


#contrast
```{r}
## Define contrasts, extract results table, and shrink the log2 fold changes

#Define your disease group first, then your control 

contrast_ <- c("SampleType", "HD", "Control")
resultsNames(HDdds)


res_table <- results(HDdds, contrast=contrast_, alpha = 0.05)


```


# summary function

```{r}
## Summarize results
summary(res_table)

```

#saving results table
```{r}

write.csv(res_table, file="results/HD_VS._control_res_table.csv")

```





#setting thresholds for DEGs
```{r}
### Set thresholds
padj.cutoff <- 0.05
lfc.cutoff <- 0.6

```


#subsetting table for DEGs values only
```{r}
HDDEGs <- res_table[which (res_table$padj < padj.cutoff & abs(res_table$log2FoldChange)>lfc.cutoff),]

write.csv(HDDEGs, file="results/HD_VS._control_DEGslfc0.6.csv")
```



```{r}
aggregate_rows <- function(df.in, agg.var){
  df.in.data <- df.in[-dim(df.in)[2]]
  df.in.data <- sapply(df.in.data, as.numeric)
  df.in.agg <- aggregate(df.in.data, list(agg.var), FUN=mean)
  rownames(df.in.agg) <- df.in.agg$Group.1
  df.in.agg <- df.in.agg[-1]   
  return(df.in.agg)
}
```


#converting ENSG of DEGs to GENE SYMBOL 
```{r}
ENSG_to_symbol <- function(df.in){
  df.ensemble.id <- sapply(rownames(df.in), function(x) strsplit(as.character(x),"\\.")[[1]][1])
  df.in <- cbind(df.ensemble.id, df.in)
  mapper <- mapIds(org.Hs.eg.db, keys=df.ensemble.id, keytype="ENSEMBL", column="SYMBOL", multiVals="first")
  mapper.df <- as.data.frame(mapper)
  mapper.df <- cbind(df.in$df.ensemble.id, mapper.df)
  names(mapper.df) <- c("df.ensemble.id", "symbol")
  df.out <- merge(df.in, mapper.df ,by="df.ensemble.id", all.x=TRUE)
  df.out <- df.out[-1]
  df.out <- df.out[ ! is.na(df.out$symbol),]
  df.out <- aggregate_rows(df.out, agg.var=df.out$symbol)
  return(df.out)
}


```

#converting and saving DEGs
```{r}
HDDEGs_symb <- ENSG_to_symbol(HDDEGs)
write.csv(HDDEGs_symb, file="results/HD_VS._control_ConvertedDEGslfc0.6.csv")

```





#getting nuclear-encoded mitochondrial DEGs only from DEGs list
```{r}
#retrieve the list file
MT_List_file <- read.csv("data/nucleargenes.csv", header=T, check.names=FALSE, row.names=1)
List_of_genes <- MT_List_file[,1]


DF <- List_of_genes [List_of_genes %in% rownames(HDDEGs_symb)]
MT_DEGs <-as.data.frame(HDDEGs_symb[DF,])

write.csv(DF, file = "results/HD_VS._control_Nuclear_encoded_DEGs_only.csv")

```



#background genes (the genes that are present in the counts file)

```{r}

tot_genes_list <- rownames(HDnormalized_counts)


```


#printing boxplot of module mitochondrial DEGs in a pdf

```{r}
module_file <-  read.csv("data/modulenmHD.csv", header=T, check.names=FALSE, row.names=1)
gene_list <- module_file$ENSG
vsd <- vst(HDdds)
c <- 1

pdf(file="results/HD_VS._control_nucleargenes_with_symbols.pdf", onefile=TRUE, paper="a4", width = 8, height = 11)
par(mfrow=c(3,3))

for (gene in gene_list){
  if (gene %in% tot_genes_list){
    colorslist = c("#0F4291", "#C40B0A")
    boxplot(t(assays(vsd[gene])[[1]])~vsd$SampleType,
            range=0, 
            las=1, 
            boxwex=.4,
            at=c(1.0 , 0.0), 
            xlab=NULL, 
            ylab=NULL,
            main=module_file[c,1],
            col=colorslist)
    stripchart(t(assays(vsd[gene])[[1]])~vsd$SampleType, 
               vertical=TRUE, 
               method='jitter', 
               add=TRUE, 
               pch=20, 
               col="black", 
               cex=1.0, 
               at=c(1.0 , 0.0))
  }
  c <- c+1
}

dev.off()
par(mfrow=c(1,1))
```


















