---
title: "AD Code"
---


```{r}
# load libraries
library(ggplot2)
library(DESeq2)
```



#Data retrieval part
```{r}
AllADData <- read.csv("data/AD_Counts.csv", header=T, check.names=FALSE, row.names=1)
AllADmeta <- read.csv("data/AD_Meta.csv", header=T, check.names = FALSE,  row.names=1)
```

#subsetting AD and Old Controls groups only
```{r}
ADmeta <- AllADmeta[(AllADmeta$SampleType2== "AD" | AllADmeta$SampleType2== "Old"),]
ADData <- AllADData[,colnames(AllADData) %in% rownames(ADmeta)]


```


#Check step
```{r}

### Check that sample names match in both files
all(colnames(ADData) %in% rownames(ADmeta))
all(colnames(ADData) == rownames(ADmeta))

```

#Saving Counts and meta files of AD and Old
```{r}
write.csv(ADData, file="data/AD_and_Old_patients_counts.csv")
write.csv(ADmeta, file="data/AD_and_Old_patients_metadata.csv")

```




#DESeq Object for comparing AD VS. Old
```{r}
## Create DESeq2Dataset object
ADdds <- DESeqDataSetFromMatrix(countData = ADData, colData = ADmeta, design = ~ SampleType1)
#View(counts(ADdds))
## Run analysis
ADdds <- DESeq(ADdds)
```



```{r}
print(ADdds)

```



#generating normalized counts
```{r}
ADdds2 <- estimateSizeFactors(ADdds)
ADnormalized_counts <- counts(ADdds2, normalized=TRUE)
write.csv(ADnormalized_counts, file="data/ADnormalized_counts.csv")

```


#contrast
```{r}
## Define contrasts, extract results table, and shrink the log2 fold changes

#Define your disease group first, then your control 

contrast_ <- c("SampleType1", "AD", "control")
resultsNames(ADdds)


res_table <- results(ADdds, contrast=contrast_, alpha = 0.05)


```


# summary function

```{r}
## Summarize results
summary(res_table)

```

#saving results table
```{r}

write.csv(res_table, file="results/AD_VS._Old_res_table.csv")



```





#setting thresholds for DEGs
```{r}
### Set thresholds
padj.cutoff <- 0.05
lfc.cutoff <- 0.6

```


#subsetting table for DEGs values only
```{r}
ADDEGs <- res_table[which (res_table$padj < padj.cutoff & abs(res_table$log2FoldChange)>lfc.cutoff),]


write.csv(ADDEGs, file="results/AD_VS._Old_DEGslfc0.6.csv")
```





#getting nuclear-encoded mitochondrial DEGs only from DEGs list
```{r}
#retrieve the genes list file
MT_List_file <- read.csv("data/nucleargenes.csv", header=T, check.names=FALSE, row.names=1)
List_of_genes <- MT_List_file[,1]



DF <- List_of_genes [List_of_genes %in% rownames(ADDEGs)]
MT_DEGs <-as.data.frame(ADDEGs[DF,])


write.csv(MT_DEGs, file = "results/AD_VS._Old_Nuclear_encoded_DEGs_only.csv")

```







#background genes (the genes that are present in the counts file)

```{r}

tot_genes_list <- rownames(ADnormalized_counts)


```


#printing boxplot of module mitochondrial DEGs in a pdf

```{r}
#reading the list of genes 
module_file <-  read.csv("data/modulenmAD.csv", header=T, check.names=FALSE, row.names=1)
gene_list <- module_file$gene

#variance stabilizing transformation of the data
vsd <- vst(ADdds)

pdf(file="results/AD_VS._Old_nucleargenes_boxplot.pdf", onefile=TRUE, paper="a4", width = 8, height = 11)
par(mfrow=c(3,3))


for (gene in gene_list){
  if (gene %in% tot_genes_list){
    colorslist = c("#C40B0A", "#0F4291")
    boxplot(t(assays(vsd[gene])[[1]])~vsd$SampleType1,
            range=0, 
            las=1, 
            boxwex=.4,
            at=c(0.0 ,0.5), 
            xlab=NULL, 
            ylab=NULL,
            main=gene,
            col=colorslist)
    stripchart(t(assays(vsd[gene])[[1]])~vsd$SampleType1,
               vertical=TRUE, 
               method='jitter', 
               add=TRUE, 
               pch=20, 
               col="black", 
               cex=1.0, 
               at=c(0.0 ,0.5))
  }
}
dev.off()
par(mfrow=c(1,1))
```


















